---
title: "Template RMarkdown for EarthCube - Short Version"
author:
- name: MyName
  email: authorone@example.com
  affiliation: Institute One
  correspondence: no
  url: http://example.com/personone
- name: MyName
  email: authortwo@example.com
  affiliation: Institute One; Insitute Two
  correspondence: yes
  url: http://example.com/persontwo
date: "2022/10/23"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
institute:
- name: Institute One
  address: 5555 Example St
  city: City
  region: Province
  country: Country
- name: Institute Two
  address: 4444 Demonstration Ave
  city: town
  region: State
  country: Country
awards:
- ShortAwardName:
    sponsor: Example.com Award Foundation
    identifier: 25002020
    name: Collaborative Award to something or other.
    url: https://example.com/awards/25002020
- NextShortAwardName:
    sponsor: Example.com Award Foundation
    identifier: 25002021
    name: Collaborative Award to something or other.
    url: https://example.com/awards/25002021

bibliography: yourbibtex.bib
keywords:
- bullet
- points
- for
- each
- keyword
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Please use the `pacman` package to load your packages:
# install.packages("pacman")
# pacman::p_load("your","required","libraries")
# pacman::p_load_gh("packages", "from", "github")
```

# Purpose
State the purpose of the notebook.




# Technical contributions

Summarize the central contributions of the notebook (libraries created, scientific analysis demonstrated, etc.), as a bullet list, with one or more bullets.


# Methodology

Quickly describe assumptions and processing steps. Include URLs as necessary.



# Results

Describe and comment on the most important results. Include images and URLs as necessary.



# Funding

Include references to awards that supported this research. 


# Keywords

Include up to 5 keywords, within the YAML header above.

# Citation

Include the recommended citation for the document.  

# Work In Progress - improvements

Use this section only if the notebook is not final.  

## Notable TODOs

Identify specific tasks here as bullets.

* todo 1
* todo 2
* todo 3

# Suggested next steps

State suggested next steps, based on results obtained in this notebook. This section is optional.

# Acknowledgements

Include any relevant acknowledgements, apart from funding (which was in section 5).


# Setup

You will need to download the Rmd, HTML and bibtex file from the earthcube repository, store it in a single directory and then set up a project. 

## Packages 
Install all the required packages in the r setup section at the head of this document. 

The code  below is an example.  


```{r}
library(jsonlite)
library(ggplot2)

x <- "hello"
y <- toJSON(x)
df <- data.frame(Country=c("ie","us","za"),Nodes=c(10,20,15))
```

## Local library import
Create a file with the source code of local packages inside the project to make sure that the folder can be reproducible in other machines and run a source command.

## Parameter definitions
Set all relevant parameters for the notebook. By convention, parameters are uppercase, while all the other variables follow R guidelines.

## Data import
Retrieve all the required data for the analysis. Include brief data descriptions, and DOIs where possible.



## Data processing and analysis
The core of the notebook is here. Split this section into subsections as required, and explain processing and analysis steps.

```{r} 
# YOUR WORK HERE!
```

Selected best practices for organizing and formatting your notebooks are included below.



# References